### 2023-10-07 latex/paper/main.tex
introduction, case study section

\pagebreak
\section{Case study}

\section{Implementation}
\subsection{Python}
Python is a versatile and high-level programming language known for its simplicity and readability. It is widely used for a variety of applications, including web development, data analysis, artificial intelligence, scientific computing, and automation. It supports multiple programming paradigms, including procedural, object-oriented, and functional programming, and its large and active community contributes to a vast ecosystem of libraries and frameworks, making it a powerful tool for a wide range of tasks. It is one of the two languages supported by Ray, alongside Java.

\subsection{Ray}
To parallelize a function with Ray, import Ray and and initialize it with \textit{ray.init()}. Then decorate the function with \textit{@ray.remote} to declare that you want to run this function remotely. Lastly, call the function with \textit{.remote()} instead of calling it normally. This remote call yields a future, a Ray object reference, that you can then fetch with \textit{ray.get}:

@@

\subsubsection{Ray clusters}
To run Ray applications on multiple nodes you must first deploy a \textbf{Ray cluster}. A Ray cluster is a set of worker nodes connected to a common Ray head node. These nodes can be physical machines, virtual machines, or containers running on a cloud infrastructure or a cluster manager like Kubernetes. Ray clusters can be fixed-size, or they may autoscale up and down according to the resources requested by applications running on the cluster.
\\\\
Each Ray cluster typically has one special node called the \textbf{head node}. This is where the main driver program is executed, which manages the cluster, schedules tasks, and communicates with the other nodes. The head node also hosts the Ray Dashboard, a web-based interface for monitoring and managing the cluster.
\\\\
The other nodes in the cluster are known as \textbf{workers}. These nodes are responsible for executing tasks and running jobs. Workers can be distributed across multiple machines and are managed by the head node. They execute user-defined Python functions and can be scaled up or down dynamically based on demand.
\\\\
A ray cluster is set up using a \textit{yaml} file, where all the details about the cluster configuration are defined. Once the file is ready, we simply launch \textit{ray up cluster.yaml} to start a ray instance on all worker nodes over ssh.
\begin{lstlisting}[caption={cluster.yaml}]
cluster_name: distmat

provider:
    type: local
    head_ip: "192.168.128.129"
    worker_ips:
    [
        "192.168.128.210",
        "192.168.128.211",
        "192.168.128.212",
        "192.168.128.213",
        "192.168.128.214",
        "192.168.128.220",
        "192.168.128.221",
        "192.168.128.222",
        "192.168.128.223",
        "192.168.128.224",
    ]

auth:
    ssh_user: <user>

upscaling_speed: 1.0
idle_timeout_minutes: 5
head_start_ray_commands:
    - ray stop
    - ulimit -c unlimited && ray start --head --port=6379 --autoscaling-config=~/ray_bootstrap_config.yaml
worker_start_ray_commands:
    - ray stop
    - ray start --address=$RAY_HEAD_IP:6379

\end{lstlisting}

\subsubsection{Cluster architecture}
/scrivere informazioni riguardo il cluster, come le specifiche tecniche e la quantita di macchine/

\pagebreak
\subsection{The Matrix class}
Before implementing the parallelized algorithms for Ray, we write the code for the serial execution in the \textbf{Matrix} class (most of the functions defined in the class are auxiliary therefore were excluded for practical purposes):

@@

\section{Result of tests}

>>>>

\subsection{Python}
Python is a versatile and high-level programming language known for its simplicity and readability. It is widely used for a variety of applications, including web development, data analysis, artificial intelligence, scientific computing, and automation. Python's strength lies in its support for multiple programming paradigms, including procedural, object-oriented, and functional programming. Its large and active community contributes to a vast ecosystem of libraries and frameworks, making it a powerful tool for a wide range of tasks.
\\\\
In the context of this paper, Python serves as the primary programming language for conducting tests and implementing linear algebra operations. Additionally, Matplotlib is utilized for graphing and visualizing the results of these tests.
\\\\
The object-oriented programming (OOP) paradigm has been employed in this research to represent matrices. This approach enhances code organization and readability while facilitating the implementation of complex linear algebra operations.
\\\\
It's worth noting that Python is one of the two languages supported by Ray, with the other being Java, making it a suitable choice for distributed computing and parallelization.

\section{Case study}

In this section, we present the case study conducted to evaluate the performance of distributed computation of linear algebra operations using Ray 2.4.0 on a cluster comprised of 10 computers with specifications provided by the University of Messina. The experiments include the caluclation of determinant, dot product, matrix inversion, and rank. For both serial and parallel execution, we conducted tests with increasing matrix sizes and repetitions to gather comprehensive performance data.
\\\\
The cluster used for our experiments is composed of ten computers with the following specifications, provided by the University of Messina:
\begin{itemize}
    \item {RAM: 8GB}
    \item {CPU: DA INSERIRE}
\end{itemize}
\pagebreak
\subsection{Execution of tests}
For our experiments, we considered the following parameters:
\begin{itemize}
    \item \textbf{Determinant calculation:} Matrices of sizes $n = 2 \dots 10$ were used, with each $n$ value being executed three times. The average execution time was recorded for analysis.
    \item \textbf{Dot product:} Matrices of sizes $n = 2 \dots 30$ were employed, with each $n$ value being executed five times.
    \item \textbf{Inverse:} Matrices of sizes $n = 2 \dots 8$ were utilized, with each $n$ value being executed three times.
    \item \textbf{Rank:} Matrices of sizes $n = 2 \dots 500$ were tested, with each $n$ value being executed five times.
\end{itemize}
\vfill
For each type of operation, the code iterates through different matrix sizes, executing the operation multiple times for each size to capture variations in performance, and recording the execution times for analysis. The results of these tests are stored in separate CSV files, organized by operation type and parameter configuration. This is achieved with the following Python script:
\pagebreak
\begin{lstlisting}[language=Python, label=main.py]
import ray
from matrix_serial import Matrix
import time
import csv

def test_dot(l, u, runs, filename):
    '''Make tests for matrices size l...u'''
    max_n = (u-l)*runs
    counter = 0
    measures = []

    print(f"running test_dot with l={l}, u={u}, runs={runs}")

    for n in range(l, u):
        sub_run = []
        sub_run.append(n) # matrix size

        for i in range(runs):
            a = Matrix.random_int(n, n, -10**8, 10**8)
            b = Matrix.random_float(n, n, -99, 99)

            t0 = time.time()
            Matrix.dot(a,b)
            t1 = time.time()

            sub_run.append(t1-t0)

            counter += 1
            print(f"run # {counter}/{max_n}: {t1-t0}")

        measures.append(sub_run)

    with open(f"{filename}_{l}_{u}_{runs}.csv", "w") as f:
        writer = csv.writer(f)
        writer.writerows(measures)

    return measures

def test_det(l, u, runs, filename):
    [...]
def test_rank(l, u, runs, filename):
    [...]
def test_inv(l, u, runs, filename):
    [...]

if __name__ == "__main__":
    if ray.is_initialized:
        ray.shutdown()
    ray.init(include_dashboard=True)

    test_det(2, 10, 3, "test_results/[[serial/parallel]]/det")
    test_dot(2, 30, 5, "test_results/[[serial/parallel]]/dot")
    test_inv(2, 8, 3, "test_results/[[serial/parallel]]/inv")
    test_rank(2, 500, 5, "test_results/[[serial/parallel]]/rank")
    \end{lstlisting}

\subsection{Graphing the results}
The code loads CSV files onto defined variables and then generates plots with error bars to visualize the performance differences between serial and parallel execution using \textit{matplotlib}. The primary goal is to assess the efficiency and scalability of these operations when executed in parallel. The resulting plots are saved as PNG files for analysis.
\\\\
A different function was defined for the plotting of rank, downsampling the data. This is because of its particular case of increasing standard deviation value, thus legibility reasons.

\begin{lstlisting}[language=Python, label=plots.py]
import csv
import matplotlib.pyplot as plt
import numpy as np

def serial_parallel_comparison(csv_name, func_name):
    # Read data from serial CSV file
    serial_data = []
    with open(f"test_results/serial/{csv_name}", "r") as f:
        reader = csv.reader(f)
        for row in reader:
            serial_data.append([float(val) for val in row])

    # Read data from parallel CSV file
    parallel_data = []
    with open(f"test_results/parallel/{csv_name}", "r") as f:
        reader = csv.reader(f)
        for row in reader:
            parallel_data.append([float(val) for val in row])

    # Extract serial data
    serial_sizes = [run[0] for run in serial_data]
    serial_means = [np.mean(runs) for runs in serial_data]
    serial_std_devs = [np.std(runs) for runs in serial_data]

    # Extract parallel data
    parallel_sizes = [run[0] for run in parallel_data]
    parallel_means = [np.mean(runs) for runs in parallel_data]
    parallel_std_devs = [np.std(runs) for runs in parallel_data]

    # Plot data
    plt.errorbar(serial_sizes, serial_means, yerr=serial_std_devs, fmt='o-', label='Serial Execution Time')
    plt.errorbar(parallel_sizes, parallel_means, yerr=parallel_std_devs, fmt='x-', label='Parallel Execution Time')
    plt.xlabel('Matrix Size (n)')
    plt.ylabel('Execution Time (s)')
    plt.title(f'Serial vs. parallel execution time ({func_name})')
    plt.grid()
    plt.legend()
    plt.savefig(csv_name.replace("csv", "png"))
    plt.show()

def rank_serial_parallel_comparison(csv_name):
    [...]
    # Downsample the data
    n = 10
    serial_sizes = serial_sizes[::n]
    serial_means = serial_means[::n]
    serial_std_devs = serial_std_devs[::n]
    parallel_sizes = parallel_sizes[::n]
    parallel_means = parallel_means[::n]
    parallel_std_devs = parallel_std_devs[::n]
    [...]

serial_parallel_comparison("det_2_10_3.csv", "determinant")
serial_parallel_comparison("dot_2_30_5.csv", "dot product")
serial_parallel_comparison("inv_2_8_3.csv", "inverse")
rank_serial_parallel_comparison("rank_2_500_5.csv")
\end{lstlisting}
\section{Implementation}
\subsection{Ray}
To parallelize a function with Ray, import Ray and and initialize it with \textit{ray.init()}. Then decorate the function with \textit{@ray.remote} to declare that you want to run this function remotely. Lastly, call the function with \textit{.remote()} instead of calling it normally. This remote call yields a future, a Ray object reference, that you can then fetch with \textit{ray.get}:

@@

\subsubsection{Ray clusters}
To run Ray applications on multiple nodes you must first deploy a \textbf{Ray cluster}. A Ray cluster is a set of worker nodes connected to a common Ray head node. These nodes can be physical machines, virtual machines, or containers running on a cloud infrastructure or a cluster manager like Kubernetes. Ray clusters can be fixed-size, or they may autoscale up and down according to the resources requested by applications running on the cluster.
\\\\
Each Ray cluster typically has one special node called the \textbf{head node}. This is where the main driver program is executed, which manages the cluster, schedules tasks, and communicates with the other nodes. The head node also hosts the Ray Dashboard, a web-based interface for monitoring and managing the cluster.
\\\\
The other nodes in the cluster are known as \textbf{workers}. These nodes are responsible for executing tasks and running jobs. Workers can be distributed across multiple machines and are managed by the head node. They execute user-defined Python functions and can be scaled up or down dynamically based on demand.
\\\\
A ray cluster is set up using a \textit{yaml} file, where all the details about the cluster configuration are defined. Once the file is ready, we simply launch \textit{ray up cluster.yaml} to start a ray instance on all worker nodes over ssh.
\begin{lstlisting}[caption={cluster.yaml}]
cluster_name: distmat
provider:
    type: local
    head_ip: "192.168.128.129"
    worker_ips:
    [
        "192.168.128.210",
        "192.168.128.211",
        "192.168.128.212",
        "192.168.128.213",
        "192.168.128.214",
        "192.168.128.220",
        "192.168.128.221",
        "192.168.128.222",
        "192.168.128.223",
        "192.168.128.224",
    ]
auth:
    ssh_user: <user>
upscaling_speed: 1.0
idle_timeout_minutes: 5
head_start_ray_commands:
    - ray stop
    - ulimit -c unlimited && ray start --head --port=6379 --autoscaling-config=~/ray_bootstrap_config.yaml
worker_start_ray_commands:
    - ray stop
    - ray start --address=$RAY_HEAD_IP:6379
\end{lstlisting}
\pagebreak
\subsection{The Matrix class}
Before implementing the parallelized algorithms for Ray, we write the code for the serial execution in the \textbf{Matrix} class (most of the functions defined in the class are auxiliary therefore were excluded for practical purposes):

@@
